---
title: "Built Environment Efforts in Woonsocket"
# subtitle: "Rhode Island Department of Health (RIDOH)" 
title-block-banner-color: "#FAFAFA"
title-block-banner: www/header.png
include-in-header:
  - text: |
      <style>
      .quarto-title-block .quarto-title-banner {
        background-position-y: center;
        height: 325px;
      }
      </style>
format: 
  html:
    page-layout: full
    backgroundcolor: "#FAFAFA"
    mermaid:
      theme: default
---

```{r, include = FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(viridis)

# ---- load: ridot bike paths
bike_paths <- read_sf("data/rigis/TRANS_Bike_Paths_RIDOT_spf_6654327448778679796.geojson") %>% 
  st_transform(4326) # re-project for mapping
# bike_paths %>% group_by(TYPE) %>% summarise(n = n()) %>% arrange(desc(n)) %>% View()
# bike_paths %>% group_by(STATUS) %>% summarise(n = n()) %>% arrange(desc(n)) %>% View()

# ---- load: bikeways and trails
# note: sf of bikes & trails (no other info)
# bikes_n_trails <- read_sf("data/rigis/Bikeways_and_Trails_-6810032319405835885.geojson") %>%
#   st_transform(4326)

# ---- load: tree equity map
tree_eq <- read_sf("data/ri_geojson/ri_tes.geojson") %>%
  st_transform(4326)
woonsocket <- tree_eq %>%
  filter(place == "Woonsocket") %>%
  reframe(geometry = st_union(geometry)) 
tree_pal <- colorBin(
  palette = c("#F5FBFF", "#8babf1"),
  domain = tree_eq$tes,
  bins = 4,
  na.color = "#E6E6E6",
  right = FALSE
)

# ---- load parks locations
parks <- read.csv("data/parks_pts.csv") %>%
  mutate(color = "#0073e6") 
parks_p <- parks %>% 
  filter(type == "Park") %>% 
  st_as_sf(coords=c("longitude","latitude"), crs=4326) 
parks_dp <- parks %>% 
  filter(type == "Dog park") %>% 
  st_as_sf(coords=c("longitude","latitude"), crs=4326)

# ---- load ripta locations
ripta <- read_sf("data/rigis/TRANS_RIPTA_Bus_Stops_2024_6929259553566897324.geojson") %>%
  st_transform(4326) %>%
  mutate(color = "#c44601")

```

::::::::::: grid
::: g-col-12
## Overview

#### A city’s built environment can provide a rich source of physical activity & active transportation for its residents

:::

::: g-col-8

Physical activity is a great way for anyone of any age to take charge of their health. Regular physical activity can result in numerous health benefits – such as decreasing risk for chronic diseases and early death. Cities that are built to support physical activity not only promote health among residents, but also enrich the community by supporting economic stimulation and social connection. 

Following the [Community Preventive Services Task Force's built environment recommendation](https://www.thecommunityguide.org/findings/physical-activity-built-environment-approaches.html), the CDC recommends focusing on everyday destinations and activity-friendly routes. Everyday destinations can be affected by increased proximity to community or neighborhood destinations, mixed land use, residential density, and parks parks access. Some measures of activity-friendly routes include: street pattern design & connectivity, pedestrian infrastructure, bicycle infrastructure, and public transit infrastructure & access.In practice this could look like sidewalks, trails, and parks that are safe and convenient for people to integrate to their daily life. For example, more than 15,000 Rhode Island commuters are pedestrians who gain daily physical activity by traveling through public streets - according to the ACS 5-Year Estimates (2009-2023).

**Together, the City of Woonsocket and advocates, including the Woonsocket HEZ & Downtown Woonsocket Collaborative, have taken measures to:**

  - **Maintain flexibility in City zoning and land use policies** to meet current needs as appropriate
  
  - **Provide regional bicycle connections and recreational opportunities** for both residents and visitors, through paths like the Blackstone River Bikeway

:::

::::: g-col-4
:::: {style="text-align: center;"}
::: {.callout-note style="border-left: solid 2px #dee2e6 !important;" appearance="minimal" icon="false"}
**Highlight #1**

Highlight #1

------------------------------------------------------------------------

**Highlight #2**

Highlight #2

------------------------------------------------------------------------

**Highlight #3**

Highlight #3
:::
::::
:::::

::: g-col-12
## History and Timeline

#### Key policies, investments, events, and other factors that have played a role advancing its built environment in a way that promotes physical activity in the community.		

```{mermaid}
%%| fig-align: center
flowchart TB
  subgraph Land Use
    B["`<b>Downtown Redevelopment Area Designation</b>
    2014`"] --> C["`<b>Mixed-Use Overlay Zoning - Downtown Main Street District</b>
    2015 (expanded 2024)`"]
    C["`<b>Mixed-Use Overlay Zoning - Downtown Main Street District</b>
    2015 (expanded 2024)`"] --> Z["`<b>Proposed zoning text & map revisions</b>
    2024`"]
  end 
  
  subgraph Pedestrian & Cyclist Networks
    F("`<b>Blackstone Valley Bikeway</b>
    2007-2018`") --> A["`<b>Bikeway Signage</b>
    2014`"]
    A["`<b>Bikeway Signage</b>
    2014`"] --> G("`<b>Truman Drive Greenway project in the City of Woonsocket</b>
    2023-2024`")
    G("`<b>Truman Drive Greenway project in the City of Woonsocket</b>
    2023-2024`") --> D["`<b>Safe Streets for All</b>
    2024`"]
    G("`<b>Truman Drive Greenway project in the City of Woonsocket</b>
    2023-2024`") --> E["`<b>Ban of electronic bikes in City parks</b>
    2024`"]
  end
```

**Pedestrian & Cyclist Networks**

  * **Blackstone Valley Bikeway (2007-2018):** a part of the Blackstone River Greenway, a 48-mile trail from Worcester, MA, to Providence, RI, has significant portions completed, but several gaps remain. While 17 miles of off-road, paved trails exist in segments, the remaining sections require funding and implementation.
  
  * **Bikweway Signage (2014)**: City partneriship with the East Coast Greenway Alliance for new directional signage around Woonsocket for the Bikepath.
  
  * **Truman Drive Greenway project (2023-2024)**: transformed a 4-lane, underutilized public road (Truman Drive) into a 2-lane road and linear park incorporating the Blackstone River Bikeway and green infrastructure to treat stormwater runoff into the nearby Blackstone River, while providing recreational, economic, and public health benefits to the city.
  
  * **Safe Streets for All** and the **ban of electronic bikes in parks (2024)**: policies to support the SSFA program, develop a Safety Action Plan, and promote safety for pedestrians in City parks by prohibiting motorized bikes.
  
**Land Use**

  * **Downtown Redevelopment Area Designation (2014)**: created a legal description of the boundaries of the redevelopment area and planning for projects within it. Redevelopment projects can include streetscapes, building renovations, new construction, flood control, water and sewer improvements, and parking lots.
  
  * **Mixed-Use Overlay Zoning (2015 - expanded 2024)**: re-zoning that made it easier for investors to propose residential and commercial (re)development in the downtown district.
    - Store-front vacancies have declined from 50% to less than 18%.  
    - Permitting multi-family residential development by-right along Main Street has resulted in approx. 600 new units of housing either recently completed or in the development pipeline, all of which helps to shape a more vibrant and walkable downtown.
    
  * **Proposed zoning text & map revisions (2024)**: as outlined in the [2024 Comprehensive Plan](https://www.woonsocketri.org/sites/g/files/vyhlif5231/f/uploads/240201_woonsocket_master_doc_-_final.pdf), the City revisited its Zoning Ordinance and Official Zoning Map to include the goals of:
    - GOAL LU-6: Providing access to open space for all residents and bringing environmental values into the City.
    - GOAL LU-7: Planning for Land Use in anticipation of improvement to the City's transportation network, and coordinating its accommodation of all users - including pedestrians, bicyclists, and automobile drivers.
    - GOAL LU-8: Encouraging the development of the Blackstone River Valley Heritage Corridor.
    - GOAL LU-9: Maintaining a Planning & Development Department that satisfactorily plans for Woonsocket's optimum physical, economic, and community development. 

:::

::: g-col-12
# Resources

## Interactive Map

```{r, echo=FALSE, message=FALSE}
leaflet(height = 650, width = "100%") %>%
  addProviderTiles(provider = "CartoDB.Positron",
                   options = providerTileOptions(minZoom = 13.5)) %>%
  addPolygons(data = woonsocket$geometry,
              opacity = 1,
              fillOpacity = 0,
              weight = 4,
              color = "#054fb9") %>%
  addPolygons(data = tree_eq, 
              color = "#054fb9",
              fillColor = ~tree_pal(tes),
              group = "Tree Equity Score",
              label = tree_eq$place,
              labelOptions = labelOptions(textsize = 14),
              popup = paste0("<strong>", tree_eq$place, "</strong><br>",
                             "<em>", tree_eq$GEOID, "</em><br>",
                             "Tree Equity Score: ", tree_eq$tes),
              fillOpacity = 0.65,
              weight = 1.2) %>%
  addPolylines(data = bike_paths, color = "#f57600",
               group = "RIDOT bike paths",
               label = str_to_title(bike_paths$PROJ_NAME),
               labelOptions = labelOptions(textsize = 14),
               opacity = 1,
               weight = 6) %>%
  addCircleMarkers(data = ripta,
                   radius = 6,
                   label = ripta$stop_name,
                   labelOptions = labelOptions(textsize = 14),
                   stroke = FALSE,
                   color = ripta$color,
                   fillOpacity = 1,
                   group = "RIPTA stops") %>%
  addCircleMarkers(data = parks_p,
                   label = parks_p$name,
                   labelOptions = labelOptions(textsize = 14),
                   stroke = FALSE,
                   color = parks_p$color,
                   fillOpacity = 1,
                   group = "Public parks") %>%
  addCircleMarkers(data = parks_dp,
                   label = parks_dp$name,
                   labelOptions = labelOptions(textsize = 14),
                   popup = parks_dp$description,
                   stroke = FALSE,
                   color = parks_dp$color,
                   fillOpacity = 1,
                   group = "Public parks") %>%
  addLegend(pal = tree_pal, 
            title = "Tree Equity Score [range: 0-100]",
            values = tree_eq$tes, 
            opacity = 1,
            group = "Tree Equity Score",
            position = "bottomright") %>%
  # add layer for woonsocket outline
  addLayersControl(position = "topright",
                   overlayGroups = c("Tree Equity Score", "Public parks", "RIDOT bike paths", "RIPTA stops"),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  htmlwidgets::onRender("
        function() {
            $('.leaflet-control-layers-overlays').prepend('<label style=\"text-align:center; font-size:16px; width:275px\"><strong>Main Features of the Built Environment in Woonsocket, RI</strong></label>');
        }
    ") %>%
  setView(lng = -71.50014,lat = 41.998603, zoom = 13.5)
```
:::

::: g-col-12
## Get Involved

* **[Attend a public hearing on Zoning](https://www.woonsocketri.gov/public-hearing-notices)**: stay up to date on public hearings for the Planning Board and Zoning Board of Review.

* **[Host an event at a Woonsocket Park](https://www.woonsocketri.gov/parks-recreation/pages/permits-process)**: apply for use of the athletic facilities or to host community events.

:::
:::::::::::
