---
title: "Built Environment Efforts in Woonsocket"
# subtitle: "Rhode Island Department of Health (RIDOH)" 
title-block-banner-color: "#FAFAFA"
title-block-banner: www/header.png
include-in-header:
  - text: |
      <style>
      .quarto-title-block .quarto-title-banner {
        background-position-y: center;
        height: 325px;
      }
      </style>
format: 
  html:
    backgroundcolor: "#FAFAFA"
---

```{r, include = FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
```

::::::::::: grid
::: g-col-12
## Overview

### *A city’s built environment can provide a rich source of physical activity & active transportation for its residents*

:::

::: g-col-8

Regular physical activity can result in numerous health benefits – such as decreasing risk for obesity and cardiovascular disease. The American Heart Association recommends policy to promote active transportation at 3 levels: the top-level of land use, the mid-level of pedestrian and bicycle networks and infrastructure, and the micro-level of street design and safety. Together, the City of Woonsocket and advocates, including the Woonsocket HEZ & Downtown Woonsocket Collaborative, have taken measures to:

  - Maintain flexibility in City zoning and land use policies to meet current needs as appropriate
  
  - Provide regional bicycle connections and recreational opportunities for both residents and visitors, through paths like the Blackstone River Bikeway

:::

::::: g-col-4
:::: {style="text-align: center;"}
::: {.callout-note style="border-left: solid 2px #dee2e6 !important;" appearance="minimal" icon="false"}
**Highlight #1**

Highlight #1

------------------------------------------------------------------------

**Highlight #2**

Highlight #2

------------------------------------------------------------------------

**Highlight #3**

Highlight #3
:::
::::
:::::

::: g-col-12
## History and Timeline

Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here

```{mermaid}
%%| fig-align: center
flowchart LR
  A[Hard edge] --> B(Round edge)
  B --> C{Decision}
  C --> D[Result one]
  C --> E[Result two]
```
:::

::: g-col-12
# Resources

## Interactive Map

```{r, echo=FALSE, message=FALSE}
# ---- load: ridot bike paths
bike_paths <- read_sf("data/rigis/TRANS_Bike_Paths_RIDOT_spf_6654327448778679796.geojson") %>% 
  st_transform(4326) # re-project for mapping
# bike_paths %>% group_by(TYPE) %>% summarise(n = n()) %>% arrange(desc(n)) %>% View()
# bike_paths %>% group_by(STATUS) %>% summarise(n = n()) %>% arrange(desc(n)) %>% View()

# ---- load: bikeways and trails
# note: sf of bikes & trails (no other info)
bikes_n_trails <- read_sf("data/rigis/Bikeways_and_Trails_-6810032319405835885.geojson") %>%
  st_transform(4326)
leaflet(width = "100%") %>%
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolylines(data = bikes_n_trails, color = "#009eb0",
               group = "Bikeways and trails per the RI state greenways plan",
               opacity = 1, 
               weight = 5) %>%
  addPolylines(data = bike_paths, color = "#029356",
               group = "RIDOT bike paths",
               label = str_to_title(bike_paths$PROJ_NAME),
               opacity = 1,
               weight = 5) %>%
  addLayersControl(position = "bottomright",
                   overlayGroups = c("Bikeways and trails per the RI State Greenways Plan",
                                     "RIDOT bike paths"),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  setView(lng = -71.5148,lat = 42.0029, zoom = 14)
```
:::

::: g-col-12
## Get Involved
:::
:::::::::::
