---
title: "Built Environment Efforts in Woonsocket"
title-block-banner-color: "#FFFFFF"
title-block-banner: www/header.png
include-in-header:
  - text: |
      <style>
      .quarto-title-block .quarto-title-banner {
        background-position-y: center;
        height: 250px;
      }
      </style>
format: 
  html:
    theme: lumen
    css: styles.css
---

```{r, include = FALSE}
library(leaflet)
```


::: grid

::: g-col-12
# Overview
:::

::: g-col-8
Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here 
:::

::: g-col-4
:::: {style="text-align: center;"}
:::{.callout-note style="border-left: solid 2px #dee2e6 !important;" appearance="minimal" icon="false"}
**Highlight #1**

Highlight #1

-------------------------------------------

**Highlight #2**

Highlight #2

-------------------------------------------

**Highlight #3**

Highlight #3
:::
::::
:::

::: g-col-12

# History and Timeline

Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here 

:::

::: g-col-12

# Interactive Map

```{r, echo=FALSE, message=FALSE}
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = -71.5148,lat = 42.0029, zoom = 14)
```


:::

::: g-col-12

# Call to Action

:::

:::