---
title: "Built Environment Efforts in Woonsocket"
# subtitle: "Rhode Island Department of Health (RIDOH)" 
title-block-banner-color: "#FAFAFA"
title-block-banner: www/header.png
include-in-header:
  - text: |
      <style>
      .quarto-title-block .quarto-title-banner {
        background-position-y: center;
        height: 325px;
      }
      </style>
format: 
  html:
    page-layout: full
    backgroundcolor: "#FAFAFA"
    mermaid:
      theme: default
---

```{r, include = FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(viridis)

# ---- load: ridot bike paths
bike_paths <- read_sf("data/rigis/TRANS_Bike_Paths_RIDOT_spf_6654327448778679796.geojson") %>% 
  st_transform(4326) # re-project for mapping
# bike_paths %>% group_by(TYPE) %>% summarise(n = n()) %>% arrange(desc(n)) %>% View()
# bike_paths %>% group_by(STATUS) %>% summarise(n = n()) %>% arrange(desc(n)) %>% View()

# ---- load: bikeways and trails
# note: sf of bikes & trails (no other info)
# bikes_n_trails <- read_sf("data/rigis/Bikeways_and_Trails_-6810032319405835885.geojson") %>%
#   st_transform(4326)

# ---- load: tree equity map
tree_eq <- read_sf("data/ri_geojson/ri_tes.geojson") %>%
  st_transform(4326)
woonsocket <- tree_eq %>%
  filter(place == "Woonsocket") %>%
  reframe(geometry = st_union(geometry)) 
tree_pal <- colorBin(
  palette = c("#F5FBFF", "#8babf1"),
  domain = tree_eq$tes,
  bins = 4,
  na.color = "#E6E6E6",
  right = FALSE
)

# ---- load parks locations
parks <- read.csv("data/parks_pts.csv") %>%
  mutate(color = "#0073e6") 
parks_p <- parks %>% 
  filter(type == "Park") %>% 
  st_as_sf(coords=c("longitude","latitude"), crs=4326) 
parks_dp <- parks %>% 
  filter(type == "Dog park") %>% 
  st_as_sf(coords=c("longitude","latitude"), crs=4326)

# ---- load ripta locations
ripta <- read_sf("data/rigis/TRANS_RIPTA_Bus_Stops_2024_6929259553566897324.geojson") %>%
  st_transform(4326) %>%
  mutate(color = "#c44601")

# ---- load seasonal events 
events <- read.csv("data/woonsocket_seasonal_events.csv") %>% 
  mutate(color = "#5928ed") %>%
  st_as_sf(coords=c("longitude","latitude"), crs=4326)
```

::::::::::: grid
::: g-col-12
## Overview

### The City of Woonsocket strives to create a built environment that provides a rich source of physical activity and active transportation for all residents

:::

::: g-col-8

Physical activity is an excellent way for people of all ages to take charge of their health. Regular physical activity results in numerous health benefits - [such as decreased risk for chronic disease and early death](https://www.cdc.gov/physical-activity/php/about/index.html). Cities that are built to support physical activity not only improve residents' health, but also enrich the community by promoting economic stimulation and social connection.

In order to create an environment conducive to physical activity, per the Community Preventive Services Task Force's [built environment recommendation](https://www.thecommunityguide.org/findings/physical-activity-built-environment-approaches.html), the CDC recommends city-planners focus on everyday destinations and activity-friendly routes. Everyday destinations can be improved by increasing their proximity to other community & neighborhood destinations, leveraging land use mechanisms, and assessing the appropriate level of residential density. As for activity-friendly routes, some measures of quality include: street pattern design & connectivity, pedestrian & bicycle infrastructure, and public transit infrastructure & access. In practice, this all might look like sidewalks, trails, and/or parks that are safe and convenient for residents to integrate into their daily routines. 

**Together, the City of Woonsocket and advocates, have taken measures to improve the built environment by:**

  - **Maintaining flexibility in City zoning and land use policies** to meet current needs as appropriate
  
  - **Providing regional bicycle connections and recreational opportunities** for both residents and visitors, through paths like the Blackstone River Bikeway

:::

::::: g-col-4
:::: {style="text-align: center;"}
::: {.callout-note style="border-left: solid 2px #dee2e6 !important;" appearance="default" icon="false"}

## Timeline Highlights

[{{< fa bicycle size=2x >}}]{style="color:#054fb9;"}

**Blackstone River Bikeway**

*2007-2018*

Totals to over 18 miles of bikeway. 

------------------------------------------------------------------------

[{{< fa store size=2x >}}]{style="color:#054fb9;"}

**Mixed-Use Overlay Zoning**

*2015 - expanded 2024*

Resulted in 600 new housing units and the decline of store-front vacancies from 50% to <18%, all contributing to a more vibrant, walkable downtown.

------------------------------------------------------------------------

[{{< fa building size=2x >}}]{style="color:#054fb9;"}

**Truman Drive Greenway project**

*2023-2024*

Incorporating the Blackstone River Bikeway - provides recreational, economic, and public health benefits.

:::
::::
:::::

::: g-col-12
## History and Timeline

### Key policies, investments, events, and other factors that have played a role advancing its built environment in a way that promotes physical activity in the community.		

```{mermaid}
%%| fig-align: center
flowchart TB
  subgraph Land Use
    B["`<b>Downtown Redevelopment Area Designation</b>
    2014`"] --> C["`<b>Mixed-Use Overlay Zoning - Downtown Main Street District</b>
    2015 (expanded 2024)`"]
    C["`<b>Mixed-Use Overlay Zoning - Downtown Main Street District</b>
    2015 (expanded 2024)`"] --> Z["`<b>Proposed zoning text & map revisions</b>
    2024`"]
  end 
  
  subgraph Pedestrian & Cyclist Networks
    F("`<b>Blackstone River Bikeway</b>
    2007-2018`") --> A["`<b>Bikeway Signage</b>
    2014`"]
    A["`<b>Bikeway Signage</b>
    2014`"] --> G("`<b>Truman Drive Greenway project in the City of Woonsocket</b>
    2023-2024`")
    G("`<b>Truman Drive Greenway project in the City of Woonsocket</b>
    2023-2024`") --> D["`<b>Safe Streets for All</b>
    2024`"]
    G("`<b>Truman Drive Greenway project in the City of Woonsocket</b>
    2023-2024`") --> E["`<b>Ban of electronic bikes in City parks</b>
    2024`"]
  end
```

**Pedestrian & Cyclist Networks**

  * **Blackstone River Bikeway (2007-2018):** a part of the Blackstone River Greenway, a 48-mile trail from Worcester, MA, to Providence, RI, has significant portions completed, but several gaps remain. While 17 miles of off-road, paved trails exist in segments, the remaining sections require funding and implementation.
  
  * **Bikweway Signage (2014)**: City partneriship with the East Coast Greenway Alliance for new directional signage around Woonsocket for the Bikepath.
  
  * **Truman Drive Greenway project (2023-2024)**: transformed a 4-lane, underutilized public road (Truman Drive) into a 2-lane road and linear park incorporating the Blackstone River Bikeway and green infrastructure to treat stormwater runoff into the nearby Blackstone River, while providing recreational, economic, and public health benefits to the city.
  
  * **Safe Streets for All** and the **ban of electronic bikes in parks (2024)**: policies to support the SSFA program, develop a Safety Action Plan, and promote safety for pedestrians in City parks by prohibiting motorized bikes.
  
**Land Use**

  * **Downtown Redevelopment Area Designation (2014)**: created a legal description of the boundaries of the redevelopment area and planning for projects within it. Redevelopment projects can include streetscapes, building renovations, new construction, flood control, water and sewer improvements, and parking lots.
  
  * **Mixed-Use Overlay Zoning (2015 - expanded 2024)**: re-zoning that made it easier for investors to propose residential and commercial (re)development in the downtown district.
    - Store-front vacancies have declined from 50% to less than 18%.  
    - Permitting multi-family residential development by-right along Main Street has resulted in approx. 600 new units of housing either recently completed or in the development pipeline, all of which helps to shape a more vibrant and walkable downtown.
    
  * **Proposed zoning text & map revisions (2024)**: as outlined in the [2024 Comprehensive Plan](https://www.woonsocketri.org/sites/g/files/vyhlif5231/f/uploads/240201_woonsocket_master_doc_-_final.pdf), the City revisited its Zoning Ordinance and Official Zoning Map to include the goals of:
    - *GOAL LU-6:* Providing access to open space for all residents and bringing environmental values into the City.
    - *GOAL LU-7:* Planning for Land Use in anticipation of improvement to the City's transportation network, and coordinating its accommodation of all users - including pedestrians, bicyclists, and automobile drivers.
    - *GOAL LU-8:* Encouraging the development of the Blackstone River Valley Heritage Corridor.
    - *GOAL LU-9:* Maintaining a Planning & Development Department that satisfactorily plans for Woonsocket's optimum physical, economic, and community development. 

:::

::: g-col-12
# Resources

## Interactive Map

```{r, echo=FALSE, message=FALSE}
leaflet(height = 650, width = "100%") %>%
  addProviderTiles(provider = "CartoDB.Positron",
                   options = providerTileOptions(minZoom = 13.5)) %>%
  addPolygons(data = woonsocket$geometry,
              opacity = 1,
              fillOpacity = 0,
              weight = 4,
              color = "#054fb9") %>%
  addPolygons(data = tree_eq, 
              color = "#054fb9",
              fillColor = ~tree_pal(tes),
              group = "Tree Equity Score",
              label = tree_eq$place,
              labelOptions = labelOptions(textsize = 14),
              popup = paste0("<strong>", tree_eq$place, "</strong><br>",
                             "<em>", tree_eq$GEOID, "</em><br>",
                             "Tree Equity Score: ", tree_eq$tes),
              fillOpacity = 0.65,
              weight = 1.2) %>%
  addPolylines(data = bike_paths, color = "#f57600",
               group = "RIDOT bike paths",
               label = str_to_title(bike_paths$PROJ_NAME),
               labelOptions = labelOptions(textsize = 14),
               opacity = 1,
               weight = 6) %>%
  addCircleMarkers(data = ripta,
                   radius = 6,
                   label = ripta$stop_name,
                   labelOptions = labelOptions(textsize = 14),
                   stroke = FALSE,
                   color = ripta$color,
                   fillOpacity = 1,
                   group = "RIPTA bus stops") %>%
  addCircleMarkers(data = parks_p,
                   label = parks_p$name,
                   labelOptions = labelOptions(textsize = 14),
                   stroke = FALSE,
                   color = parks_p$color,
                   fillOpacity = 1,
                   group = "Public parks") %>%
  addCircleMarkers(data = parks_dp,
                   label = parks_dp$name,
                   labelOptions = labelOptions(textsize = 14),
                   popup = paste0("<strong>", parks_dp$name, "</strong><br>",
                                  parks_dp$description),
                   stroke = FALSE,
                   color = parks_dp$color,
                   fillOpacity = 1,
                   group = "Public parks") %>%
  addCircleMarkers(data = events,
                   label = events$event_name,
                   labelOptions = labelOptions(textsize = 14),
                   popup = paste0("<strong>", events$event_name, "</strong><br>",
                                  events$description),
                   stroke = FALSE,
                   color = events$color,
                   fillOpacity = 0.5,
                   group = "Seasonal events") %>%
  addLegend(pal = tree_pal, 
            title = "Tree Equity Score [range: 0-100]",
            values = tree_eq$tes, 
            opacity = 1,
            group = "Tree Equity Score",
            position = "bottomright") %>%
  # add layer for woonsocket outline
  addLayersControl(position = "topright",
                   overlayGroups = c("Tree Equity Score", "Public parks", "RIDOT bike paths", "RIPTA bus stops", "Seasonal events"),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  htmlwidgets::onRender("
        function() {
            $('.leaflet-control-layers-overlays').prepend('<label style=\"text-align:center; font-size:16px; width:275px\"><strong>Main Features of the Built Environment in Woonsocket, RI</strong></label>');
        }
    ") %>%
  setView(lng = -71.50014,lat = 41.998603, zoom = 13.5)
```
:::

::: g-col-12
## Get Involved

* **[Attend a public hearing on Zoning](https://www.woonsocketri.gov/public-hearing-notices)**: stay up to date on public hearings for the Planning Board and Zoning Board of Review.

* **[Host an event at a Woonsocket Park](https://www.woonsocketri.gov/parks-recreation/pages/permits-process)**: apply for use of the athletic facilities or to host community events.

:::
:::::::::::
