---
title: "Built Environment Efforts in Woonsocket"
subtitle: "Rhode Island Department of Health (RIDOH)" 
title-block-banner-color: "#FFFFFF"
title-block-banner: www/header_edited.png
include-in-header:
  - text: |
      <style>
      .quarto-title-block .quarto-title-banner {
        background-position-y: center;
        height: 285px;
      }
      </style>
format: 
  html:
    mermaid: 
      theme: forest
    theme: lumen
    backgroundcolor: "#FAFAFA"
    css: styles.css
---

```{r, include = FALSE}
library(leaflet)
library(ggplot2)
```

::::::::::: grid
::: g-col-12
# Overview

## How the built environment can support access to physical activity & active transportation
:::

::: g-col-8
Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here 

Fake plot

```{r, message=FALSE, echo=FALSE}
ggplot(data = mtcars,
       aes(x = gear, y = carb)) + 
  geom_point()
```


:::

::::: g-col-4
:::: {style="text-align: center;"}
::: {.callout-note style="border-left: solid 2px #dee2e6 !important;" appearance="minimal" icon="false"}
**Highlight #1**

Highlight #1

------------------------------------------------------------------------

**Highlight #2**

Highlight #2

------------------------------------------------------------------------

**Highlight #3**

Highlight #3
:::
::::
:::::

::: g-col-12
## History and Timeline

Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here Text here

```{mermaid}
%%| fig-align: center
flowchart LR
  A[Hard edge] --> B(Round edge)
  B --> C{Decision}
  C --> D[Result one]
  C --> E[Result two]
```
:::

::: g-col-12
# Resources

## Interactive Map

```{r, echo=FALSE, message=FALSE}
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = -71.5148,lat = 42.0029, zoom = 14)
```
:::

::: g-col-12
## Get Involved
:::
:::::::::::
